---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
activity <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

Total numbers of steps taken per day
```{r}
steps_d <- with(activity, tapply(steps, date, sum, na.rm = TRUE))
```

Histogram of total steps taken per day
```{r}
library(ggplot2)
ggplot()+geom_histogram(aes(steps_d), bins=5, color="White") + ggtitle("Histogram of steps taken per day") + xlab("Steps")
```

Mean value of the total number of steps per day
```{r}
mean(steps_d)
```

Median value of the total number of steps per day
```{r}
median(steps_d)
```

## What is the average daily activity pattern?

Time series plot of the average number of steps taken in 5-minute interval
```{r}
steps_daily <- with(activity, tapply(steps, interval, mean, na.rm = TRUE))
library(ggplot2)
ggplot(data.frame(day = as.numeric(names(steps_daily)), step = steps_daily)) + geom_line(aes(day, step))
```

The 5-minute interval contains maximum number of steps on average across all the days
```{r}
names(which.max(steps_daily))
```

## Imputing missing values

Total Number of rows contains at least one NA
```{r}
sum(!complete.cases(activity))
```

Replace NA to the mean steps for that day.
```{r}
activity_imp <- activity
for(i in seq_len(nrow(activity_imp))){
        if(is.na(activity_imp[i, "steps"])){
                activity_imp[i, "steps"] <- steps_d[activity_imp[i, "date"]]
        }
}
```

Total numbers of steps taken per day
```{r}
steps_d_imp <- with(activity_imp, tapply(steps, date, sum, na.rm = TRUE))
```

Histogram of total steps taken per day
```{r}
#hist(steps_d_imp, main = ("Histogram of steps taken per day"), xlab = "Steps")
library(ggplot2)
ggplot()+geom_histogram(aes(steps_d_imp), bins=5, color="White") + ggtitle("Histogram of steps taken per day") + xlab("Steps")
```

Mean value of the total number of steps per day
```{r}
mean(steps_d_imp)
```

Median value of the total number of steps per day
```{r}
median(steps_d_imp)
```
Histogram, mean, and median total number of steps taken per day do not changed after imputing missing data to the mean value.

## Are there differences in activity patterns between weekdays and weekends?
A new factor variable in the dataset with two levels - "weekday" and "weekend"
Add the factor to a new colume "week" into the activity_imp
```{r}
weekend <- c("Saturday", "Sunday")
activity_imp$week <- factor((weekdays(as.Date(activity_imp$date)) %in% weekend), levels = c(FALSE, TRUE), labels = c("Weekday", "Weekend"))
```


```{r}
steps_dailyw <- with(activity_imp, aggregate(steps, list(interval, week), mean, na.rm = TRUE))

library(ggplot2)
ggplot(steps_dailyw, aes(Group.1, x)) + geom_line(aes(group = Group.2)) + facet_wrap(vars(Group.2), nrow = 2) + xlab("Interval") + ylab("Number of steps")
```